Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-12-17 main.tex
Discussion of the findings

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of the Findings}

The four school studies, the experiment at the surgery and the two case-studies
provide a first broad picture of how a robot-based remediation to handwriting is
accepted by the children and practitioners, and what are the outcomes that can
be expected.

In total, more than 80 children have interacted with the system, for a total
duration of more than 38 hours in multiple experimental configurations
(individual interaction \vs pairs \vs groups; at school, in the lab, at a
surgery; short interactions \vs long, repeated interactions; 5 to 8 years old
children).

@@

We believe that this research provides a novel perspective on educative robots
at several levels. We show that:

@@

As we show through extensive experimental validation, these claims are not just
words: we have effectively deployed robots on the field, with children suffering
actual handwriting impairments. Our results are promising in terms of sustaining
the interest and attention of children on otherwise boring writing exercises,
and we evidence handwriting improvements.

It is however still early to quantify the lasting effects of this remediation:
handwriting is a complex cognitive skill, that builds on many individual and
social factors. Self-confidence is one of them. Our approach that endows the
child with the role of a teacher who can help a robot: we expect it may help
some children to recover self-esteem and self-confidence. However the
experiments that we have conducted so far do not allow us to confirm this
hypothesis, and more research need to be conducted in this direction.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary of the Findings and Discussion}

The four school studies, the experiment at the surgery and the two case-studies
provide a first broad picture of how a robot-based remediation to handwriting is
accepted by the children and practitioners, and what outcomes can be expected.

In total, more than 80 children have interacted with the system, for a total
duration of more than 38 hours, in multiple experimental configurations
(individual interaction \vs pairs \vs groups; at school, in the lab, at a
surgery; short interactions \vs long, repeated interactions; 5 to 8 years old
children). We summarize hereafter the main findings from these experiments and
discuss some of the critical points of our approach.

\paragraph{Technical Assessment} We set ourselves the challenge of developing an
autonomous robotic system able to perform handwriting tasks with children.
Over the course of the different studies and experiments, thousands of
letters have been indeed generated and drawn on the tactile tablet by the robot, in
response to thousands of demonstrations from children. Most of the experiments have been
carried out in the wild, at six different locations in total. As reported in the
technical studies section, the system performed generally well: the children had no
issues interacting with the robot, the robot experienced relatively few crashes,
and those crashes did not significantly impaired the interactions. The system
has furthermore proved robust enough to conduct several hour-long experiments.
However, we also found that the level of expertise required to deploy and
operate our system still makes the presence of an experimenter mandatory at all
time. Non-trivial technical developments may be required to reach a level of
usability suitable for a broader, non-expert, audience.

\paragraph{Acceptance} When we started the project, the mere acceptance of the
robot in the educative and/or medical context of handwriting remediation was not
obvious. Expectedly, the attitude of practitioners toward the robot indeed
varies a lot at individual levels. The process that led us to select the schools
and surgeries for the experiments is interesting in that regard. The very first
school we contacted was a local public school. The head teacher was positive
about the study, but did ask for official agreement, which never happened. As we
learnt several months later, the request was still laying on the desk of a
regional education officer who happens not to be too keen on novel technologies.
Conversely, the three schools where we did conduct the first studies are all
private schools, generally open to novel learning methods. The teachers we met
during these experiments certainly had questions about the rationale and
end-purpose of the project, but were confident about its educative and social
value.

Similarly with the occupational therapists: while Thomas' therapist did
voluntarily contacted us (after hearing about the project at the radio) and
readily involved herself in the design of the experiment with Thomas, we had
less positive feedback from other practitioners: in one instance, we contacted
ourselves a local group of school psychologists and occupational therapists to
present the project: after the meeting, the opinions were definitively mixed,
with some therapists willing to conduct actual experiments with their children,
and others not quite as enthusiastic.

Lastly (and as expected), the reaction of the children to the robot was good:
they enjoyed interacting with the system, and, as we have shown, they actually
committed to their teacher role. The teaching situation was well accepted: while
we had initial doubts about how believable a learner the robot would be, the
children did not appear to show any specific issues with it. In particular,
technical choices like having the robot to only gesture writing on the tablet
instead of actually \emph{physically} writing with a pen, were not raised as
issues.

\paragraph{Engagement} We measured the engagement of the children by three
different means: the number of demonstrations they provided to the robot, the
amount of qualitative feedback they gave to the robot (in the \emph{Surgery}
experiment), and to what extend they were focusing on the task (measure of the
with-me-ness). Independently of the experimental setting, it appears that the
children engage easily into the interaction. More interestingly, we show in the
\emph{Surgery} experiment that they generally take on the role of a teacher
easily, that they act this role seriously (and not only playfully), and that
they can assume this role over an extended period of time. Designing a system
that keeps children engaged over several hours is especially important for
handwriting remediation, and we show in the case-studies that even children as
young as 5.5 years old like Thomas were able to do so. This seems to indicate
that our approach could be relevant for a broad range of ages.

Another finding relates to the Protégé effect: as seen with Vincent's case-study
(with the parents emphasizing how Vincent was aware he had to be consistent in
his writing to help the robot, or with him sending us a mail several months
later to know how the robot was doing with its writing) or with Thomas's
case-study (when Thomas realizes that the robot ``does not understand'' when his
demonstrations are not legible enough, and consequently quickly improves his
own writing to better help the robot), our system does seem to elicit a Protégé
effect that not only help the children to remain engaged into their teaching
over several weeks, but also positively impact their learning process (Vincent
strives to write in a consistent manner, as does Thomas). This supports {\it a
posteriori} our choice to build an interaction situation based on the
\emph{learning by teaching} paradigm.

\paragraph{Remediation Efficacy} Vincent's case study did provide us with
initial material to evidence handwriting improvements (Figure~\ref{fig:stimuli})
and the study with Thomas provide further data, both quantitative
(Figure~\ref{Thomas_progress}) and qualitative (feedback from the therapist that
points how Thomas is much better at drawing consistent shapes in a repeated manner,
as well as reflecting on his own performance by training several times before
actually sending a demonstration to the robot).

We must however remain cautious here as for the actual role of the system: while
the children were indeed the one deciding what to teach to the robot and the
robot was autonomously learning and responding, the role of the adults (the
experimenters or the occupational therapists) should not be underestimated.
Beyond the normal explanations on how to operate the tablet and how to interact
with the robot, the adults played the role of a facilitator in every of the
studies by prompting the children to comment on the robot's performance,
suggesting possible corrections, or proposing to try another letter/number/word.
This facilitation not only compensates for the possible shortcomings of the interaction,
but is also a fundamental part of the learning process itself. In that respect,
our robot is essentially a tool that creates a favourable learning situation for
the child, and where the adult (be it a teacher or a therapist) keeps its entire
educative role.

@@

We believe that this research provides a novel perspective on educative robots
at several levels. We have shown that:

@@

As we show through extensive experimental validation, these claims are not just
words: we have effectively deployed robots on the field, with children suffering
actual handwriting impairments. Our results are promising in terms of sustaining
the interest and attention of children on otherwise repetitive writing exercises,
and we evidence handwriting improvements.

It is however still early to quantify the lasting effects of this remediation:
handwriting is a complex cognitive skill, that builds on many individual and
social factors. Self-confidence is one of them. Our approach endows the child
with the role of a teacher who can help a robot: we expect it may as well help
some children to recover self-esteem and self-confidence by putting them in a
positive, gratifying role. The experiments that we have conducted so far do not
allow us to confirm this hypothesis yet, and more research will have to be
conducted in this direction.
